Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-07-15 doc/report/msc-report.tex
report: added references.
report: speling corrections and other minorities.

HDW is a system whose goal is to create a high performance large scale data
warehouse based on MapReduce and Bigtable. \cite{hdw}

@@

In chapter 2 general data warehous concepts are introduced; it also introduces
new techniques that the reader might not know.

@@

% CONCEPTS IN DATA WAREHOUSING, by clark de jean
\chapter{Concepts}

@@

Classic data warehouses use relational databases as storage. A star or snowflake
schema would be used to modell the dimensions.

@@

% import.
The Zohmg import component concists of a framework which enables the user to
write import programs. The import component handles half of the MapReduce
pattern, the reduce part; it also adds an interface for relaying the imported
data to the storage component. The import programs written by the user include
the other half of the MapReduce pattern, namely the map part, and uses the
interface for storing the imported data.

@@

% fit in existing eco system.
The system needs to co-exist and integrate seamlessly with the already existing
hardware and software eco-system at Last.fm. Thus, building on top of already
existing software is a natural choice.

@@

Furthermore the requirements on the Zohmg system were discussed; the four key
requirements are: ease of use, speed, scaling, and fitting into the existing
hardware and software eco-system.

@@

Once a thourough understanding of the problem scope and its theoretical solutian
the design were entered. Finally a short implementation phase was executed.

@@

Also a lot of related and surronding, possibly useful, concepts were studied;
although some of these never made it to the final solution.
\cite{pig,mapreducemerge}

@@

These databases were explored and tested, but most of them were in a very early
phase of development and hard to even install. The experience was that HBase was
the most mature of them. Maturity was the compound experience with the software:
community, codebase, documentation, stability etc. HBase is also tightly coupled
with Hadoop which is used in production at Last.fm.

@@

After familiarity with the problem and its theoretical solution was reached,
the design phase was entered. During this phase a grand design for the system
was layed out.

The initial design was ambitious; it contained, amongst other things, components
to import, store, and exporte data, a query language and sub-components for
query analyzation. The intended tasks for these components were to identify the
most common queries and optimization. These features are likely advantageous in
a large data store but it is possible to do without them.

@@

The goal of HDFS is to store large data sets while being able to handle
hardware failure and being deployed in heterogenous hardware and software
eco systems. Additionally, the Hadoop and HDFS interaction is designed with
the notion that it is cheaper to move the computation, MapReduce programs,
to the data than the other way around.

@@

All measures are integers. Sum is the only supported aggregation function. The
data store does not support reach-like data (i.e. number of unique visitors)
since it is not summable.

@@

In the first section the implementation of project creation, configuration, and
setup, will be presented. The second section features the data import
implementation; it will present example data as well as an example user map
function to import this. The third section presents the implementation of data
interchange between Zohmg and other software, such as Hadoop, Dumbo, and HBase,
but also throughout the Zohmg system. The third section also presents the
implementation details of the Zohmg system's exporting compontent.

@@

The data set files reside in the project's \texttt{config} directory. The data
sets are configured using YAML\footnote{YAML Ain't a Markup Language}, a
human-readable data serializaton format. Any human-readable configure file
method could have been chosen. YAML was choosen because it is is very simple
for humans to read and edit, and likewise for machines. The Python module
PyYAML was used to implement the YAML configuration reading part. \cite{pyyaml}

@@

A standard MapReduce map function emits key-value pairs. The Zohmg usermap
function, instead, emits triples (tuples with three values), which consist of a
time-stamp, a hash table of dimensions and their respective values, and a hash
table with measurements (values). In effect, the dimensions and the time-stamp
specify a point in n-space and the measurements represent the observations made
at that point.

@@

For every line of the log file (the value argument to the map function), a point
in n-space (the dimensions and the time-stamp) is emitted together with a count
of pageviews and bytes; one pageview per request and the number of bytes served
in the request. It is possible to add logic here, perhaps only counting requests
from useragents that are not web spiders which resulted in a HTTP 200 status
code.

@@

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the directory \texttt{/data/weblogs/2009/04/20}
on HDFS. The emitted output is duly collected and projections are carefully
stored in the data store.

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form \\
\texttt{\{"column-family:qualifier" : value\}}.

@@

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
framework, which includes both middleware and application parts. The middleware
applications are responsible for dispatching requests to the corresponding
application or file. WSGI middleware receives a WSGI request and then performs
logic upon this request, before passing the request to a WSGI application.
\cite{paste,definitive_guide_to_pylons} Since the main goal was not to build a
middleware framework from scratch, an existing such was used.

@@

The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009. The results from the scan is served to the user as
JSONP\footnote{JSON with padding: JSON extension which adds name of callback
function as an argument to the function itself.}.

@@

This chapter's first section has presented the implementation of the Zohmg
system. The concept of Zohmg project's and management of thesame were
introduced; as well as how these are configured. The second section presented
how to import data with Zohmg, giving examples of data and the user map
function. In the third, and final, section Zohmg's data interchange was
featured. Both the internal data interchange, which is used together with
Hadoop, Dumbo, and HBase, and the external data interchange, which is used by
the user to query Zohmg.

@@

The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009.

@@

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map function (in the MapReduce sense) and letting Zohmg run that map function
over the data set. The user's map function is called \textit{usermap function}
from here on. In order to let Zohmg know what output to expect from the mapper,
the user also defines the dimensions and units of his data. After all, the user
most likely knows his own data better than anyone else.

>>>>

HDW is a system whose goal is to create a high performance large scale data
warehouse based on MapReduce and BigTable. \cite{hdw}

@@

In chapter 2 general data warehouse concepts are introduced; it also introduces
new techniques that the reader might not know.

@@

\chapter{Concepts}

@@

Classic data warehouses use relational databases as storage. A star or snowflake
schema would be used to model the dimensions.

@@

% import.
The Zohmg import component consists of a framework which enables the user to
write import programs. The import component handles half of the MapReduce
pattern, the reduce part; it also adds an interface for relaying the imported
data to the storage component. The import programs written by the user include
the other half of the MapReduce pattern, namely the map part, and uses the
interface for storing the imported data.

@@

% fit in existing eco system.
The system needs to coexist and integrate seamlessly with the already existing
hardware and software eco system at Last.fm. Thus, building on top of already
existing software is a natural choice.

@@

Furthermore the requirements on the Zohmg system were discussed; the four key
requirements are: ease of use, speed, scaling, and fitting into the existing
hardware and software eco system.

@@

Once a thorough understanding of the problem scope and its theoretical solution
the design were entered. Finally a short implementation phase was executed.

@@

Also a lot of related and surrounding, possibly useful, concepts were studied;
although some of these never made it to the final solution.
\cite{piglatin,mapreducemerge}

@@

These databases were explored and tested, but most of them were in a very early
phase of development and hard to even install. The experience was that HBase was
the most mature of them. Maturity was the compound experience with the software:
community, code base, documentation, stability etc. HBase is also tightly coupled
with Hadoop which is used in production at Last.fm.

@@

After familiarity with the problem and its theoretical solution was reached,
the design phase was entered. During this phase a grand design for the system
was laid out.

The initial design was ambitious; it contained, amongst other things, components
to import, store, and export data, a query language and sub-components for
query analysis. The intended tasks for these components were to identify the
most common queries and optimization. These features are likely advantageous in
a large data store but it is possible to do without them.

@@

The goal of HDFS is to store large data sets while being able to handle
hardware failure and being deployed in heterogeneous hardware and software
eco systems. Additionally, the Hadoop and HDFS interaction is designed with
the notion that it is cheaper to move the computation, MapReduce programs,
to the data than the other way around.

@@

All measures are integers. Sum is the only supported aggregation function. The
data store does not support reach-like data (i.e. number of unique visitors)
since it is not sumable.

@@

In the first section the implementation of project creation, configuration, and
setup, will be presented. The second section features the data import
implementation; it will present example data as well as an example user map
function to import this. The third section presents the implementation of data
interchange between Zohmg and other software, such as Hadoop, Dumbo, and HBase,
but also throughout the Zohmg system. The third section also presents the
implementation details of the Zohmg system's exporting component.

@@

The data set files reside in the project's \texttt{config} directory. The data
sets are configured using YAML\footnote{YAML Ain't a Markup Language}, a
human-readable data serialization format. Any human-readable configure file
method could have been chosen. YAML was chosen because it is is very simple
for humans to read and edit, and likewise for machines. The Python module
PyYAML was used to implement the YAML configuration reading part. \cite{pyyaml}

@@

A standard MapReduce map function emits key-value pairs. The Zohmg user map
function, instead, emits triples (tuples with three values), which consist of a
time-stamp, a hash table of dimensions and their respective values, and a hash
table with measurements (values). In effect, the dimensions and the time-stamp
specify a point in n-space and the measurements represent the observations made
at that point.

@@

For every line of the log file (the value argument to the map function), a point
in n-space (the dimensions and the time-stamp) is emitted together with a count
of page views and bytes; one page view per request and the number of bytes served
in the request. It is possible to add logic here, perhaps only counting requests
from user agents that are not web spiders which resulted in a HTTP 200 status
code.

@@

This will instruct Zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the directory \texttt{/data/weblogs/2009/04/20}
on HDFS. The emitted output is duly collected and projections are carefully
stored in the data store.

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what row key,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form \\
\texttt{\{"column-family:qualifier" : value\}}.

@@

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
framework, which includes both middleware and application parts. The middle
ware applications are responsible for dispatching requests to the corresponding
application or file. WSGI middleware receives a WSGI request and then performs
logic upon this request, before passing the request to a WSGI application.
\cite{paste,definitive_guide_to_pylons} Since the main goal was not to build a
middleware framework from scratch, an existing such was used.

@@

The above query would return the page views for SE between 1 January
2009 and 4 April 2009. The results from the scan is served to the user as
JSONP\footnote{JSON with padding: JSON extension which adds name of callback
function as an argument to the function itself.}.

@@

This chapter's first section has presented the implementation of the Zohmg
system. The concept of Zohmg project's and management of the same were
introduced; as well as how these are configured. The second section presented
how to import data with Zohmg, giving examples of data and the user map
function. In the third, and final, section Zohmg's data interchange was
featured. Both the internal data interchange, which is used together with
Hadoop, Dumbo, and HBase, and the external data interchange, which is used by
the user to query Zohmg.

@@

The above query would return the page views for SE between 1 January
2009 and 4 April 2009.

@@

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map function (in the MapReduce sense) and letting Zohmg run that map function
over the data set. The user's map function is called \textit{user map function}
from here on. In order to let Zohmg know what output to expect from the mapper,
the user also defines the dimensions and units of his data. After all, the user
most likely knows his own data better than anyone else.
